### 2019-06-30 DNP_Meetings/abstract_corrected.tex
update

\documentclass[a4paper]{article}
%\usepackage{simplemargins}

%\usepackage[square]{natbib}
%\usepackage{amsmath}
%\usepackage{amsfonts}
%\usepackage{amssymb}
%\usepackage{graphicx}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
%\usepackage[urw-garamond]{mathdesign}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{tgbonum}
\usepackage[left=4cm, right=4cm, top=1cm]{geometry}

\begin{document}
\pagenumbering{gobble}

\Large
 \begin{center}
\textbf{First Cross Section Results of D(e,e'p)n at Very High Recoil Momenta\\}

\hspace{10pt}

% Author names and affiliations
\large
Carlos Yero \\

\hspace{10pt}

\small
for the Hall C Collaboration

\hspace{10pt}

\small
(Dated: July 01, 2019) \\

\hspace{10pt}

\small  
Florida International University\\

\end{center}

\hspace{10pt}

\large
\begin{center}
ABSTRACT
\end{center}
\normalsize

\noindent Preliminary D(e,e'p)n electro-disintegration cross sections at $Q^{2} = 4.25$ GeV$^{2}$ with recoil
momenta up to 900 MeV/c will be presented. The experiment ran for a total of 6 days of beam time in April 2018
at Jefferson Lab in Hall C and it seeks to study the short range structure of the deuteron by probing its high
momentum components beyond 500 MeV/c, where currently no data exists. The experiment was part of a group of Hall C
experiemnts that commissioned the new Hall C Super High Momentum Spectrometer (SHMS). At the selected kinematics,
Meson Exchange Currents (MEC) and Isobar Configurations (IC) are suppressed. Final State Interactions (FSI) have also
been suppressed by choosing a kinematic region where the neutron recoil angle is between 35 and 45 degrees with respect
to the momentum transfer. This suppression was seen in a previous D(e,e'p)n experiment (Boeglin et al. Phys Rev Lett. 2011)
and is also predicted in modern theoretical calculations (W. Boeglin \& M. Sargsian Int.J.Mod.Phys. 2015). In this region,
the Plane Wave Impulse Approximation (PWIA) dominates and comparisions between measured and predicted cross sections are
sensitive to the deuteron momentum distributions. Comparisons between data and different calculations with different NN
potentials will be shown.

\end{document}