### 2020-02-05 MY_PUBLICATIONS/PRL/prl_files_WBCorrections/body.tex
updated

\indent A 10.6 GeV electron beam was incident on a 10 cm long liquid deuterium target (LD2). The scattered electron and knocked-out proton were detected in coincidence
by the new Super High Momentum Spectrometer (SHMS) and the High Momentum Spectrometer (HMS), respectively. The recoiling neutron was reconstructed using energy-momentum conservation. The missing momentum is defined as 
$\vec{p_{r}} = \vec{q} - \vec{p_{f}}$ and the missing  energy as $E_{m} = \omega - T_{p} - T_{r}$ where $\vec{p_{f}}$ is the final proton momentum, $\vec{q}$ is the 3-momentum transfer and $(T_{p}$ and $T_{r})$ are the
final proton and neutron kinetic energies respectively. The beam currents delivered by the accelerator ranged between 45-60 $\mu$A and the beam was rastered over a 2x2 mm$^{2}$ area to reduce the effects of localized boiling on the cryogenic targets (hydrogen and deuterium).\\
\indent Both Hall C spectrometers have similar standard detector packages, each with four scintillator planes\cite{hodo_techreport} used for triggering, a pair of drift chambers\cite{dc_techreport} used for tracking, and a calorimeter\cite{Mkrtchyan_2013} and gas \u{C}erenkov \cite{Li_Wenliang_mthesis,ngc_techreport} used for electron identification.
The trigger setup for this experiment required a minimum of 3/4 hodoscope planes for data readout in each spectrometer. Due to the absence of significant background and the negligible
accidental trigger rates at the higher missing momentum settings, the use of additional particle identification (PID) measures was found to have little to no effect on the final cross section.\\
\indent We measured three central missing momentum settings: $p_{r}=80,580$ and $750$ MeV/c. At each of these settings, the electron arm (SHMS) was fixed and the proton arm (HMS) was rotated from smaller to larger angles corresponding to the
the lower and higher missing momentum settings, respectively. At these kinematic settings, the 3-momentum transfer covered a range of $2.4\lesssim|\vec{q}|\lesssim3.2$ GeV/c which is more than twice the highest neutron recoil momentum ($p_{r}$)
measured on this experiment. As a result, most of the virtual photon momentum is transferred to the proton which scatters at angles relative to $\vec{q}$ in the range $0.4^{o}\lesssim \theta_{pq}\lesssim21.4^{o}$.
At these forward angles and large momentum transferred to the proton, the  process where the neutron is struck by the virtual photon is suppressed.\\
\indent Hydrogen elastic $^{1}H(e,e'p)$ data was also taken at kinematics close to the deuteron $p_{r}$=80 MeV setting for cross-checks with the spectrometer acceptance model using the  Hall C Monte Carlo
simulation program, SIMC. Additional $^{1}H(e,e'p)$ data were also taken at three other kinematic settings that covered the SHMS momentum acceptance range for the deuteron and were used for spectrometer optics optimization, 
momentum calibration and the determination of the spectrometer offsets and kinematic uncertainties\cite{cyero_specKinUnc,cyero_SHMSOptics}.\\
\indent Identical event selection criteria were used for the hydrogen and deuteron data. The criteria were determined by making standard cuts on the spectrometer momentum fraction ($\delta$) to select a region in which the reconstruction optics
is well known,  a cut to restrict the HMS solid angle acceptance to events that passed directly through the collimator and not by re-scattering from the collimator edges, a missing
energy cut (peak $\sim$ 2.22 MeV for the deuteron) to select true $^{2}H(e,e'p)n$ coincidences, a coincidence time cut to select true coincidence events and not accidentals,  a PID cut on the
SHMS calorimeter normalized total track energy to select electrons and not other sources of background (mostly pions) and a cut on the reconstructed HMS and SHMS reaction vertices to select events that 
originated from the same reaction vertex at the target.\\
\indent The experimental data yields for both hydrogen and deuteron data were normalized by the total charge and corrected for various inefficiencies. For $^{2}H(e,e'p)n$ the corrections
were as follows: tracking efficiencies ($98.9 \%$-HMS, $96.4 \%$-SHMS), total live time ($92.3 \%$), proton loss due to nuclear interactions in the HMS ($4.7 \%$)\cite{cyero_pAbs} and
target boiling factors ($4.2 \%$)\cite{cyero_tgtBoil}. The values in parentheses were averaged over all data sets. \\
\indent For $^{1}H(e,e'p)$, the corrected data yield was compared to SIMC calculations using J. Arrington's proton form factor parametrization\cite{PhysRevC.69.022201} to check the spectrometer acceptance
model. The ratio of data to simulation yield was determined to be $97.6\pm0.3 \%$ (statistical uncertainty only). For $^{2}H(e,e'p)n$, the low missing momentum data ($p_{r}=80$ MeV/c) were compared to the Hall A data (See Fig. \ref{fig:fig1}).
The good agreement gives us confidence on the measurements made at higher missing momentum settings for which no previous data exist. \\
\indent The systematic uncertainties of the measured cross sections were determined from 
normalization\footnote{Conservative estimates on the systematic uncertainties of the total live and charge were made.
Determination of systematics on these quantities is a work in progress. (Private communication with D. Mack)} and kinematic uncertainties in the beam energy and spectrometer angle/momentum settings. The individual
contributions from normalization uncertainties were determined to be: tracking efficiencies ($0.40 \%$-HMS, $0.59 \%$-SHMS ), target boiling ($0.39 \%$), total live time ($3.0 \%$) and total charge ($2.0\%$)
for an overall normalization uncertainty added in quadrature of $3.7 \%$. \\
\indent The systematic uncertainties due to our limited knowledge of the beam energy and spectrometer angle/momentum settings were determined point-to-point in ($\theta_{nq}, p_{r}$) bins for each data set independently, and added in quadrature for overlapping $p_{r}$ bins
of different data sets. For $\theta_{nq}= 35^{o}, 45^{o} and 75^{o}$ (presented on this Letter) the overall kinematic uncertainty varied up to 6.5$\%$ for $p_{r}\leq1.01$ GeV/c.
The overall systematic uncertainty in the cross section was determined by the quadrature sum of the normalization and kinematic uncertainties. This result was then added in quadrature
to the statistical uncertainty(25-30$\%$ on average) to obtain the final uncertainty in the cross section. \\
\indent The data were radiatively corrected for each bin in ($\theta_{nq}, p_{r}$) by multiplying the measured cross sections by the ratio of the calculated particle yield excluding and including radiative effects. The Monte-Carlo code SIMC was used for these calculations with the Deuteron Model by J.M.Laget including FSI~\cite{LAGET2005}.
For each bin in ($\theta_{nq}, p_{r}$), the averaged $^{2}H(e,e'p)n$ kinematics has been calculated and used in the bin centering correction factor defined as:
$f_{bc} \equiv \sigma_{avg.kin} / \bar{\sigma}$, where $\sigma_{avg.kin}$ is the cross section calculated at the averaged kinematics and $\bar{\sigma}$ is the cross section averaged
over the kinematic bin.\\
\indent Both experimental and theoretical reduced cross sections were extracted from the measured and calculated cross sections for each data set independently and subsequently averaged
for overlapping bins in $p_{r}$. The reduced cross sections are defined as follows:
\begin{equation}
\sigma_{red} \equiv \frac{\sigma_{exp(th)}}{Kf_{rec}\sigma_{cc1}}
\label{eq:1}
\end{equation}
where $\sigma_{exp(th)}$ is the 5-fold experimental (or theoretical) differential cross section $\frac{d^{5}\sigma}{d\omega d\Omega_{e} d\Omega_{p}}$, $K$ is a kinematical factor, $f_{rec}$ is the recoil factor  and $\sigma_{cc1}$ is the de Forest\cite{DEFOREST1983} electron-proton off-shell cross section calculated using the form factor parametrization of Ref.\cite{PhysRevC.69.022201}.
Within the PWIA, $\sigma_{red}$ corresponds to the proton momentum distribution inside the deuteron. \\
\indent Figure \ref{fig:fig1} shows the extracted experimental and theoretical reduced cross sections as a function of neutron recoil momentum
$p_{r}$ for three recoil angle settings at
\onecolumngrid
\begin{center}
\begin{figure}[hb!]
\includegraphics[scale=0.46]{prl_plots/PRL_plot1.pdf}
\caption{The reduced cross sections $\sigma_{red}(p_{r})$ as a function of neutron recoil momentum $p_{r}$ are shown in (a)-(c) for recoil angles $\theta_{nq}=35^{o}, 45^{o}$ and $75^{o}$, respectively,
with a bin width of $\pm 5^{o}$. The data is compared to the previous Hall A experiment (cyan) results\cite{PhysRevLett.107.262501} as well as the theoretical reduced cross sections using the Paris(blue),
AV18(green) and CD-Bonn(magenta) NN potentials}
\label{fig:fig1}
\end{figure}
\end{center}
\twocolumngrid
\noindent $Q^{2}=4.5\pm0.5$ (GeV/c)$^{2}$. The results from the previous Hall A experiment\cite{PhysRevLett.107.262501} at a $Q^{2}=3.5\pm0.25$ (GeV/c)$^{2}$ are plotted as well (cyan). The data are compared to theoretical reduced
cross sections using the charge-dependent Bonn (CD-Bonn)\cite{PhysRevC.63.024001}, Argonne $v_{18}$ (AV18)\cite{PhysRevC.51.38} and Paris\cite{PhysRevC.21.861} NN-potentials. The theoretical calculations
for the CD-Bonn (magenta) and AV18 (green) potentials were performed by M. Sargsian\cite{PhysRevC.82.014612} and those for the Paris potential  (blue) were  by J.M. Laget\cite{LAGET2005}.\\
\indent For all recoil angles shown in Fig. \hyperref[fig:fig1]{1} at recoil momenta $p_{r}\leq$300 MeV/c the cross sections are well reproduced  by  all models when FSI are included.
The agreement at $p_{r}\leq300$ MeV/c can be understood from the fact that this region corresponds to the long-range part of the NN potential where the One Pion Exchange Potential (OPEP)
is well known and common to all modern potentials. \\
\indent Beyond $p_{r}\sim$300 MeV/c at $\theta_{nq}=35^{o}$ and $45^{o}$ [Figs. \hyperref[fig:fig1]{1(a), 1(b)}], the Paris and
AV18 models significantly differ from the CD-Bonn. In this region, the Paris/AV18 cross sections are dominated by the PWIA and within good agreement of each other up $p_{r}\sim700$ MeV/c. The CD-Bonn cross sections in contrast are generally smaller than the Paris/AV18 in this region.  In addition for $\theta_{nq}=35^{o}$ they are dominated by the PWIA up to $p_{r}\sim800$ MeV/c [Fig. \hyperref[fig:fig1]{1(a)}]  while for $\theta_{nq}=45^{o}$  FSI start to contribute already above 600 MeV/c [Fig. \hyperref[fig:fig1]{1(b)}].
The main difference between the CD-Bonn and Paris/AV18 models is the use of Feynman amplitudes in covariant (original) form by the Bonn group as opposed to local (static) approximations of these amplitudes used by Paris/AV18 groups
to describe the NN potential. The effect of these local approximations on the NN potential are shown in Fig. 2 of Ref. \cite{PhysRevC.63.024001}. \\
\indent At $\theta_{nq}=75^{o}$ [Fig. \hyperref[fig:fig1]{1(c)}] and $p_{r}>300$ MeV/c, FSI become the dominant contributions to the cross sections for all models which exhibit a similar
behaviour (smaller falloff) that overshadows any possibility of extracting the momentum distributions.\\
\indent To quantify the discrepancy observed between data and theory in Fig. \ref{fig:fig1}, the ratio of the experimental and theoretical reduced cross sections ($\sigma_{red}$) to the
deuteron momentum distribution  calculated using the CD-Bonn potential ($\sigma^{\text{CD-Bonn PWIA}}_{red}$)\cite{PhysRevC.63.024001} is shown in Fig. \ref{fig:fig2}.
\begin{figure}[h!]
\includegraphics[scale=0.5]{./prl_plots/PRL_plot2.pdf}
\caption{The ratio R($p_{r}$) = $\sigma_{red}/\sigma^{{\text{CD-Bonn PWIA}}}_{red}$ is shown in (a)-(c) for $\theta_{nq}=35^{o}, 45^{o}$ and $75^{o}$, respectively, each with a bin width of $\pm 5^{o}$.
The dashed reference (magenta) line refers to CD-Bonn PWIA calculation (or momentum distribution) by which the data and all models are divided.}
\label{fig:fig2}
\end{figure}
\indent For $\theta_{nq}=35^{o}$ and $45^{o}$ [Figs. \hyperref[fig:fig2]{2(a),(b)}] the data are best described by the CD-Bonn PWIA calculation for recoil momenta up to $p_{r}\sim700$ MeV/c and $\sim600$ MeV/c, respectively.
Furthermore, the agreement between the Halls A and C data solidifies the Hall A approach of selecting a kinematic region where recoil angles are small and FSI are reduced.
At larger recoil momenta where the ratio is $R>1$ and increasing.  At $\theta_{nq}=35^{o}$ FSI start to dominate for missing moment above typically above 800 MeV/c for the CD-Bonn calculation
while the other models predict still relatively small FSI below 900 MeV/c. At $\theta_{nq}=45^{o}$ the FSI dominance starts earlier for all models above 800 MeV/c and for the CD-Bonn based
calculation above 600 MeV/c. \\
\indent Overall it is interesting to note that none of the calculations can reproduce the measured $p_r$ dependence above 600 MeV/c in a region where FSI are still relatively small (<30\%).  This behavior of the data is new and 
and additional data in this kinematic region are necessary to improve the statistics. \\
\indent At $\theta_{nq}=75^{o}$ [Fig. \hyperref[fig:fig2]{2(c)}], FSI are small below $p_{r}\sim150$ MeV/c, but do not exactly cancel the PWIA/FSI interference term in the scattering amplitude which results in a small dip in this region in agreement with the data.
At $p_{r}>300$ MeV/c, the data was statistically limited as our focus was on the smaller recoil angles. The Hall A data, however, shows a reasonable agreement with the FSI from all models which gives us confidence in our understanding
of FSI at the smaller recoil angles. \\
\indent This experiment extended the previous Hall A cross section measurements on the $^{2}H(e,e'p)n$ reaction to 
very high neutron recoil momenta at kinematics where FSI were expected to be small and the cross section was dominated by PWIA and sensitive to the
short range part of the deuteron wavefunction. The experimental reduced cross sections were extracted and found to be in good agreement with the Hall A data.
Furthermore, the CD-Bonn model was found to be significantly different than the Paris or AV18 models and was able to partially describe the data over a larger range in $p_{r}$.
At higher missing momentum, however, all models were unable to describe the data. \\
\indent We acknowledge the outstanding support of the staff of the Accelerator and Physics Divisions at Jefferson Lab
as well as the entire Hall C staff, technicians, graduate students and users who took shifts or contributed
to the equipment for the Hall C upgrade making all four commissioning experiments possible.

%-------------------------------