### 2020-02-12 MY_PUBLICATIONS/PRL/draft_v4_HCsuggestions/author_listv2.tex
updated

\author{C.~Yero$^{1}$}
\author{W.U.~Boeglin$^{1}$}
\author{M.K.~Jones$^{2}$}

\affiliation{$^{1}$Florida International University, University Park, Florida 33199, USA}
\affiliation{$^{2}$Thomas Jefferson National Accelerator Facility, Newport News, Virginia 23606, USA}
\collaboration{for the Hall C Collaboration}
\noaffiliation