### 2020-02-12 MY_PUBLICATIONS/PRL/draft_v4_HCsuggestions/author_listv2.tex
updated

%Author List from Spring 2018 Hall C Commissioning + SIDIS Runs
%Comment / Uncomment accordingly
\author{C.~Yero$^{1}$}       %Main author for E12-10-003 experiment
\author{D.~Abrams}           %Commissioning ONLY
\author{Z.~Ahmed}            %Commissioning ONLY
\author{B.~Aljawrneh}        %Commissioning ONLY
\author{S.~Alsalmi}          %Commissioning ONLY
\author{R.~Ambrose}          %Commissioning + SIDIS
%\author{D.~Androic}         %Color Transparency ONLY
\author{W.~Armstrong}        %Commissioning ONLY
\author{A.~Asaturyan}        %Commissioning ONLY
\author{K.~Assumin-Gyimah}   %Commissioning ONLY
\author{A.~Bandari}          %Commissioning + SIDIS
\author{J.~Bane}             %Deuteron Experiment ONLY
\author{S.~Basnet}           %Commissioning + SIDIS
\author{V.~Berdnikov}        %Commissioning + SIDIS
\author{H.~Bhatt}            %Commissioning + SIDIS
\author{D.~Bhetuwal}         %Commissioning + SIDIS
\author{D.~Biswas}           %Commissioning + SIDIS
\author{W.~Boeglin}          %Commissioning ONLY (Deuteron Spokesperson)
\author{P.~Bosted}           %Commissioning ONLY (SIDIS Spokesperson)
\author{E.~Brash}            %Commissioning ONLY
\author{M.~Bukhari}          %Commissioning ONLY
%\author{M.~Carmignotto}     %NO MINIMUM SHIFTS
\author{H.~Chen}             %Commissioning ONLY
\author{J.P.~Chen}           %Commissioning ONLY
\author{M.~Chen}             %Commissioning ONLY
\author{E.~Christy}          %Commissioning ONLY
\author{S.C.~Dusa}           %Commissioning + SIDIS               
\author{K.~Craycraft}        %Commissioning + SIDIS
%\author{S.~Danagoulian}     %NO MINIMUM SHIFTS
%\author{D.~Day}             %ONLY DEC. 2017 Spec. Commissioning ?????? (should be included?)
\author{M.~Diefenthaler}     %Commissioning + SIDIS
\author{M.~Dlamini}          %Commissioning ONLY
\author{J.~Dunne}            %Commissioning ONLY
\author{B.~Duran}            %Commissioning + SIDIS
\author{D.~Dutta}            %Commissioning + SIDIS  (CT Spokesperson)
\author{R.~Ent}              %NO MINIMUM SHIFTS JLab Physics Division Leader (SIDIS Co-Spokesperson) (should be included)??????
\author{R.~Evans}            %Commissioning ONLY
\author{H.~Fenker}           %Commissioning + SIDIS
%\author{I.~Fernando}        %NO MINIMUM SHIFTS
\author{N.~Fomin}            %Commissioning ONLY
\author{E.~Fuchey}           %Commissioning ONLY
\author{D.~Gaskell}          %Commissioning + SIDIS
\author{T.N.~Gautam}         %Commissioning ONLY
\author{C.A.~Gayoso}         %Commissioning + SIDIS
\author{F.A.~Gonzalez}       %Commissioning + SIDIS
\author{O.~Hansen}           %Commissioning ONLY
\author{F.~Hauenstein}       %Commissioning ONLY
\author{T.~Horn}             %Commissioning ONLY
\author{G.~Huber}            %Commissioning ONLY
\author{M.K.~Jones}          %Commissioning + SIDIS
\author{S.~Joosten}          %Commissioning ONLY
\author{M.L.~Kabir}          %Commissioning + SIDIS           
%\author{N.~Kalantarians}    %NO MINIMUM SHIFTS
\author{A.~Karki}            %Commissioning + SIDIS  
\author{C.~Keppel}           %NO MINIMUM SHIFTS Hall A/C Leader (F2 Spokesperson) 
\author{A.~Khanal}           %Commissioning ONLY
\author{P.~King}             %Commissioning ONLY
\author{E.~Kinney}           %Commissioning ONLY (SIDIS Co-Spokesperson)
%\author{H.S.~Ko}            %Color Transparency ONLY
%\author{M.~Khol}            %NO MINIMUM SHIFTS
\author{N.~Lashley-Colthirst}      %Commissioning ONLY
\author{S.~Li}               %Commissioning ONLY
\author{W.~Li}               %Commissioning ONLY
\author{A.H.~Liyanage}       %Commissioning + SIDIS
\author{D.~Mack}             %Commissioning + SIDIS
\author{S.~Malace}           %Commissioning + SIDIS (F2 Spokesperson)
%\author{P.~Markowitz}       %NO MINIMUM SHIFTS
\author{J.~Matter}           %Commissioning + SIDIS
\author{D.~Meekins}          %NO MINIMUM SHIFTS (TARGET WORK DURING COMMISSIONING+SIDIS. (INCLUDED AS AUTHOR!)
\author{R.~Michaels}         %Commissioning + SIDIS
\author{H.~Mkrtchyan}        %Commissioning ONLY
\author{S.J.~Nazeer}         %Commissioning ONLY
%\author{S.~Nanda}           %Spec. commissioning only Dec. 14-17 shifts
\author{D.~Nguyen}           %Commissioning ONLY (Took 10 shifts) 
\author{G.~Niculescu}        %Took 11 shifts (Why is NOT included in Holly's list) ??????
\author{M.~Niculescu}        %Took 11 shifts (Why is NOT included in Holly's list) ??????
\author{N.~Nuruzzaman}       %Commissioning ONLY
\author{B.~Pandey}           %NO MINIMUM SHIFTS (WORKED ON HMS DCs for 1 year. INCLUDED AS AUTHOR!)
\author{S.~Park}             %Commissioning + SIDIS
\author{E.~Pooser}           %Commissioning + SIDIS
%\author{A.~Puckett}         %NO MINIMUM SHIFTS
\author{M.~Rehfuss}          %Commissioning ONLY
\author{J.~Reinhold}         %Commissioning ONLY
%\author{N.~Santiesteban}     %Color Transparency ONLY
\author{B.~Sawatzky}          %Color Transparency ONLY (Worked on DAQ/Detector. INCLUDED AS AUTHOR!)
\author{G.~Smith}             %Commissioning ONLY
\author{A.~Sun}               %Commissioning + SIDIS
\author{H.~Szumila-Vance}     %Commissioning + SIDIS
%\author{A.~Tadepalli}        %F2 ONLY
\author{V.~Tadevosyan}        %Commissioning + SIDIS
%\author{R.~Trotta}           %NO MINIMUM SHIFTS
\author{A.V.~Hernandez}       %Commissioning ONLY
\author{S.A.~Wood}            %NO MINIMUM SHIFTS (Hall C Deputy + Worked on Hall C 12 GeV software + hardware work) (INCLUDED AS AUTHOR!)
\author{Jinlong~Zhang}        %Commissioning ONLY How to distinguish between J. Zhang and J.Zhang (below) ?????? Middle names?             
\author{Jixie~Zhang}          %NO MINIMUM SHIFTS

%Spokespeople
%\author{W.U.~Boeglin$^{1}$}
%\author{M.K.~Jones$^{2}$}

\affiliation{$^{1}$Florida International University, University Park, Florida 33199, USA}

%\affiliation{$^{2}$Thomas Jefferson National Accelerator Facility, Newport News, Virginia 23606, USA}

\collaboration{for the Hall C Collaboration}
\noaffiliation