### 2020-04-10 UTILITIES_CODES/latex_files/supp.tex
updated

\title{\small \textbf{Supplemental Material: $^{2}$H$(e,e'p)n$ Cross Section Measurements \\for Neutron Recoil Momenta up to $p_{r}\sim1.01$ GeV/c at $Q^{2}=4.5\pm0.5$ (GeV/c)$^{2}$ }}

\documentclass[12pt]{article}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\begin{document}
\maketitle

\section{\small Spectrometer Acceptance}
In Hall C, the spectrometer acceptance is defined by the variables reconstructed at the target: $\delta, Y_{\mathrm{tar}}, X'_{\mathrm{tar}}$ and $Y'_{\mathrm{tar}}$, where

\end{document}