### 2020-07-27 MY_PUBLICATIONS/PRL/FINAL_DRAFT/MARKED_PRL/authorlist-rev.tex
added directory with final PRL draft

%Author List from Spring 2018 Hall C Commissioning + SIDIS Runs
%Comment / Uncomment accordingly
\author{C.~Yero$^{1}$}       %Main author for E12-10-003 experiment
\author{D.~Abrams$^{2}$}           %Commissioning ONLY
\author{Z.~Ahmed$^{3}$}            %Commissioning ONLY
\author{A.~Ahmidouch$^{4}$}        % . . .----> HOLLY
\author{B.~Aljawrneh$^{4}$}        %Commissioning ONLY
\author{S.~Alsalmi$^{5}$}          %Commissioning ONLY
\author{R.~Ambrose$^{3}$}          %Commissioning + SIDIS
%\author{D.~Androic}         %Color Transparency ONLY

\author{W.~Armstrong$^{6}$}        %Commissioning ONLY
\author{A.~Asaturyan$^{7}$}        %Commissioning ONLY
\author{K.~Assumin-Gyimah$^{8}$}   %Commissioning ONLY
\author{C.~Ayerbe Gayoso$^{9}$}    %Commissioning + SIDIS
\author{A.~Bandari$^{9}$}          %Commissioning + SIDIS
\author{J.~Bane$^{10}$}             %Deuteron Experiment ONLY
\author{S.~Basnet$^{3}$}           %Commissioning + SIDIS
\author{V.V.~Berdnikov$^{11}$}        %Commissioning + SIDIS  (tell HOLLY name change)
\author{J.~Bericic$^{15}$}          %SHMS Optics Contributor  HOLLY
\author{H.~Bhatt$^{8}$}            %Commissioning + SIDIS
\author{D.~Bhetuwal$^{8}$}         %Commissioning + SIDIS
\author{D.~Biswas$^{12}$}           %Commissioning + SIDIS
\author{W.U.~Boeglin$^{1}$}          %Commissioning ONLY (Deuteron Spokesperson)
\author{P.~Bosted$^{9}$}           %Commissioning ONLY (SIDIS Spokesperson)
\author{E.~Brash$^{13}$}            %Commissioning ONLY
\author{M.H.S.~Bukhari$^{14}$}          %Commissioning ONLY
%\author{M.~Carmignotto}     %NO MINIMUM SHIFTS
\author{H.~Chen$^{2}$}             %Commissioning ONLY
\author{J.P.~Chen$^{15}$}           %Commissioning ONLY
\author{M.~Chen$^{2}$}             %Commissioning ONLY
\author{M.E.~Christy$^{12}$}          %Commissioning ONLY
\author{S.~Covrig$^{15}$}           %Commissioning + SIDIS               
\author{K.~Craycraft$^{10}$}        %Commissioning + SIDIS
\author{S.~Danagoulian$^{4}$}            %NO MINIMUM SHIFTS. Contributed to detectros
\author{D.~Day$^{2}$}             %ONLY DEC. 2017 Spec. Commissioning Contributed to Noble Gas Detector
\author{M.~Diefenthaler$^{15}$}     %Commissioning + SIDIS
\author{M.~Dlamini$^{17}$}          %Commissioning ONLY
\author{J.~Dunne$^{8}$}
%Commissioning ONLY
\author{B.~Duran$^{16}$}            %Commissioning + SIDIS
\author{D.~Dutta$^{8}$}            %Commissioning + SIDIS  (CT Spokesperson)
\author{R.~Ent$^{15}$}              %NO MINIMUM SHIFTS JLab Physics Division Leader (SIDIS Co-Spokesperson) (should be included)??????
\author{R.~Evans$^{3}$}            %Commissioning ONLY
\author{H.~Fenker$^{15}$}           %Commissioning + SIDIS
%\author{I.~Fernando}        %NO MINIMUM SHIFTS
\author{N.~Fomin$^{10}$}            %Commissioning ONLY
\author{E.~Fuchey$^{18}$}           %Commissioning ONLY
\author{D.~Gaskell$^{15}$}          %Commissioning + SIDIS
\author{T.N.~Gautam$^{12}$}         %Commissioning ONLY
\author{F.A.~Gonzalez$^{19}$}       %Commissioning + SIDIS
\author{J.O.~Hansen$^{15}$}           %Commissioning ONLY
\author{F.~Hauenstein$^{20}$}       %Commissioning ONLY
\author{A.V.~Hernandez$^{11}$}       %Commissioning ONLY
\author{T.~Horn$^{11}$}             %Commissioning ONLY
\author{G.M.~Huber$^{3}$}            %Commissioning ONLY
\homepage[ORCiD: ]{http://orcid.org/0000-0002-5658-1065}
\author{M.K.~Jones$^{15}$}          %Commissioning + SIDIS
\author{S.~Joosten$^{6}$}          %Commissioning ONLY
\author{M.L.~Kabir$^{8}$}          %Commissioning + SIDIS           
%\author{N.~Kalantarians}    %NO MINIMUM SHIFTS
\author{A.~Karki$^{8}$}            %Commissioning + SIDIS  
\author{C.E.~Keppel$^{15}$}           %NO MINIMUM SHIFTS Hall A/C Leader (F2 Spokesperson) 
\author{A.~Khanal$^{1}$}           %Commissioning ONLY
\author{P.~King$^{17}$}             %Commissioning ONLY
\author{E.~Kinney$^{21}$}           %Commissioning ONLY (SIDIS Co-Spokesperson)
%\author{H.S.~Ko}            %Color Transparency ONLY
%\author{M.~Khol}            %NO MINIMUM SHIFTS
\author{N.~Lashley-Colthirst$^{12}$}      %Commissioning ONLY
\author{S.~Li$^{22}$}               %Commissioning ONLY
\author{W.B.~Li$^{9}$}               %Commissioning ONLY  tell HOLLY name change
\author{A.H.~Liyanage$^{12}$}       %Commissioning + SIDIS
\author{D.J.~Mack$^{15}$}             %Commissioning + SIDIS
\author{S.P.~Malace$^{15}$}           %Commissioning + SIDIS (F2 Spokesperson)
%\author{P.~Markowitz}       %NO MINIMUM SHIFTS
\author{J.~Matter$^{2}$}           %Commissioning + SIDIS
\author{D.~Meekins$^{15}$}          %NO MINIMUM SHIFTS (TARGET WORK DURING COMMISSIONING+SIDIS. (INCLUDED AS AUTHOR!)
\author{R.~Michaels$^{15}$}         %Commissioning + SIDIS
\author{A.~Mkrtchyan$^{7}$}        %Commissioning ONLY     HOLLY
\author{H.~Mkrtchyan$^{7}$}        %Commissioning ONLY
\author{S.J.~Nazeer$^{12}$}         %Commissioning ONLY
\author{S.~Nanda$^{8}$}           %Spec. commissioning only Dec. 14-17 shifts
\author{G.~Niculescu$^{23}$}        %Took 11 shifts (Why is NOT included in Holly's list) ??????
\author{M.~Niculescu$^{23}$}        %Took 11 shifts (Why is NOT included in Holly's list) ??????
\author{D.~Nguyen$^{2}$}           %Commissioning ONLY (Took 10 shifts) 
\author{N.~Nuruzzaman$^{24}$}       %Commissioning ONLY
\author{B.~Pandey$^{12}$}           %NO MINIMUM SHIFTS (WORKED ON HMS DCs for 1 year. INCLUDED AS AUTHOR!)
\author{S.~Park$^{19}$}             %Commissioning + SIDIS
\author{C.F.~Perdrisat$^{9}$}            %College of William & Mary   HOLLY
\author{E.~Pooser$^{15}$}           %Commissioning + SIDIS
%\author{A.~Puckett}         %NO MINIMUM SHIFTS
\author{M.~Rehfuss$^{16}$}          %Commissioning ONLY
\author{J.~Reinhold$^{1}$}         %Commissioning ONLY
%\author{N.~Santiesteban}     %Color Transparency ONLY
\author{B.~Sawatzky$^{15}$}          %Color Transparency ONLY (Worked on DAQ/Detector. INCLUDED AS AUTHOR!)
\author{G.~Smith$^{15}$}             %Commissioning ONLY
\author{A.~Sun$^{25}$}               %Commissioning + SIDIS
\author{H.~Szumila-Vance$^{15}$}     %Commissioning + SIDIS
%\author{A.~Tadepalli}        %F2 ONLY
\author{V.~Tadevosyan$^{7}$}        %Commissioning + SIDIS
%\author{R.~Trotta}           %NO MINIMUM SHIFTS
\author{S.A.~Wood$^{15}$}            %NO MINIMUM SHIFTS (Hall C Deputy + Worked on Hall C 12 GeV software + hardware work) (INCLUDED AS AUTHOR!)
\author{J.~Zhang$^{19}$}        %Commissioning ONLY 
%\author{Jixie~Zhang}          %NO MINIMUM SHIFTS

\affiliation{$^{1}$Florida International University, University Park, Florida 33199, USA}
\affiliation{$^{2}$University of Virginia, Charlottesville, Virginia 22903, USA}
\affiliation{$^{3}$University of Regina, Regina, Saskatchewan S4S 0A2, Canada}
\affiliation{$^{4}$North Carolina Agricultural and Technical State University, Greensboro, North Carolina 27411, USA}
\affiliation{$^{5}$Kent State University, Kent, Ohio 44240, USA}
\affiliation{$^{6}$Argonne National Laboratory, Lemont, Illinois 60439, USA}   %fixed address for Whit and Sylvester
\affiliation{$^{7}$A.I. Alikhanyan National Science Laboratory (Yerevan Physics Institute), 2 Alikhanian Brothers Street, 0036, Yerevan, Armenia}
\affiliation{$^{8}$Mississippi State University, Mississippi State, Mississippi 39762, USA}
\affiliation{$^{9}$The College of William \& Mary, Williamsburg, Virginia 23185, USA}
\affiliation{$^{10}$University of Tennessee, Knoxville, Tennessee 37996, USA}
\affiliation{$^{11}$Catholic University of America, Washington, DC 20064, USA}
\affiliation{$^{12}$Hampton University, Hampton, Virginia 23669, USA}
\affiliation{$^{13}$Christopher Newport University, Newport News, Virginia 23606, USA}
\affiliation{$^{14}$Jazan University, Jazan 45142, Saudi Arabia}
\affiliation{$^{15}$Thomas Jefferson National Accelerator Facility, Newport News, Virginia 23606, USA}
\affiliation{$^{16}$Temple University, Philadelphia, Pennsylvania 19122, USA}
\affiliation{$^{17}$Ohio University, Athens, Ohio 45701, USA}
\affiliation{$^{18}$University of Connecticut, Storrs, Connecticut 06269, USA}
\affiliation{$^{19}$Stony Brook University, Stony Brook, New York 11794, USA}
\affiliation{$^{20}$Old Dominion University, Norfolk, Virginia 23529, USA}
\affiliation{$^{21}$University of Colorado Boulder, Boulder, Colorado 80309, USA}
\affiliation{$^{22}$University of New Hampshire, Durham, New Hampshire 03824, USA}
\affiliation{$^{23}$James Madison University, Harrisonburg, Virginia 22807, USA}
\affiliation{$^{24}$Rutgers University, New Brunswick, New Jersey 08854, USA}
\affiliation{$^{25}$Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA}