### 2020-08-18 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/authorlist.tex
updated article submission docs

%Author List from Spring 2018 Hall C Commissioning + SIDIS Runs
\author{C.~Yero$^{1,15}$}       %Main author for E12-10-003 experiment
\homepage[ORCiD: ]{https://orcid.org/0000-0003-2822-7373}
\email[email: ]{cyero@jlab.org}
\author{D.~Abrams$^{2}$}           
\author{Z.~Ahmed$^{3}$}            
\author{A.~Ahmidouch$^{4}$}        
\author{B.~Aljawrneh$^{4}$}        
\author{S.~Alsalmi$^{5}$}          
\author{R.~Ambrose$^{3}$}          
\author{W.~Armstrong$^{6}$}        
\author{A.~Asaturyan$^{7}$}        
\author{K.~Assumin-Gyimah$^{8}$}   
\author{C.~Ayerbe Gayoso$^{9}$}    
\author{A.~Bandari$^{9}$}          
\author{J.~Bane$^{10}$}            
\author{S.~Basnet$^{3}$}           
\author{V.V.~Berdnikov$^{11}$}     
\author{J.~Bericic$^{15}$}         
\author{H.~Bhatt$^{8}$}            
\author{D.~Bhetuwal$^{8}$}         
\author{D.~Biswas$^{12}$}          
\author{W.U.~Boeglin$^{1}$}        
\author{P.~Bosted$^{9}$}           
\author{E.~Brash$^{13}$}           
\author{M.H.S.~Bukhari$^{14}$}     
\author{H.~Chen$^{2}$}             
\author{J.P.~Chen$^{15}$}          
\author{M.~Chen$^{2}$}             
\author{M.E.~Christy$^{12}$}       
\author{S.~Covrig$^{15}$}          
\author{K.~Craycraft$^{10}$}       
\author{S.~Danagoulian$^{4}$}      
\author{D.~Day$^{2}$}             
\author{M.~Diefenthaler$^{15}$}    
\author{M.~Dlamini$^{17}$}         
\author{J.~Dunne$^{8}$}
\author{B.~Duran$^{16}$}          
\author{D.~Dutta$^{8}$}          
\author{R.~Ent$^{15}$}            
\author{R.~Evans$^{3}$}           
\author{H.~Fenker$^{15}$}         
\author{N.~Fomin$^{10}$}          
\author{E.~Fuchey$^{18}$}         
\author{D.~Gaskell$^{15}$}        
\author{T.N.~Gautam$^{12}$}       
\author{F.A.~Gonzalez$^{19}$}     
\author{J.O.~Hansen$^{15}$}       
\author{F.~Hauenstein$^{20}$}     
\author{A.V.~Hernandez$^{11}$}    
\author{T.~Horn$^{11}$}           
\author{G.M.~Huber$^{3}$}         
\homepage[ORCiD: ]{http://orcid.org/0000-0002-5658-1065}
\author{M.K.~Jones$^{15}$}         
\author{S.~Joosten$^{6}$}          
\author{M.L.~Kabir$^{8}$}            
\author{A.~Karki$^{8}$}           
\author{C.E.~Keppel$^{15}$}          
\author{A.~Khanal$^{1}$}       
\author{P.~King$^{17}$}             
\author{E.~Kinney$^{21}$}           
\author{N.~Lashley-Colthirst$^{12}$}     
\author{S.~Li$^{22}$}               
\author{W.B.~Li$^{9}$}              
\author{A.H.~Liyanage$^{12}$}       
\author{D.J.~Mack$^{15}$}           
\author{S.P.~Malace$^{15}$}         
\author{J.~Matter$^{2}$}           
\author{D.~Meekins$^{15}$}          
\author{R.~Michaels$^{15}$}         
\author{A.~Mkrtchyan$^{7}$}        
\author{H.~Mkrtchyan$^{7}$}        
\author{S.J.~Nazeer$^{12}$}         
\author{S.~Nanda$^{8}$}           
\author{G.~Niculescu$^{23}$}        
\author{M.~Niculescu$^{23}$}        
\author{D.~Nguyen$^{2}$}           
\author{N.~Nuruzzaman$^{24}$}       
\author{B.~Pandey$^{12}$}           
\author{S.~Park$^{19}$}             
\author{C.F.~Perdrisat$^{9}$}       
\author{E.~Pooser$^{15}$}           
\author{M.~Rehfuss$^{16}$}          
\author{J.~Reinhold$^{1}$}         
\author{B.~Sawatzky$^{15}$}         
\author{G.R.~Smith$^{15}$}          
\author{A.~Sun$^{25}$}              
\author{H.~Szumila-Vance$^{15}$}    
\author{V.~Tadevosyan$^{7}$}        
\author{S.A.~Wood$^{15}$}           
\author{J.~Zhang$^{19}$}

\collaboration{for the Hall C Collaboration}
\affiliation{$^{1}$Florida International University, University Park, Florida 33199, USA}
\affiliation{$^{2}$University of Virginia, Charlottesville, Virginia 22903, USA}
\affiliation{$^{3}$University of Regina, Regina, Saskatchewan S4S 0A2, Canada}
\affiliation{$^{4}$North Carolina Agricultural and Technical State University, Greensboro, North Carolina 27411, USA}
\affiliation{$^{5}$Kent State University, Kent, Ohio 44240, USA}
\affiliation{$^{6}$Argonne National Laboratory, Lemont, Illinois 60439, USA}  
\affiliation{$^{7}$A.I. Alikhanyan National Science Laboratory (Yerevan Physics Institute), 2 Alikhanian Brothers Street, 0036, Yerevan, Armenia}
\affiliation{$^{8}$Mississippi State University, Mississippi State, Mississippi 39762, USA}
\affiliation{$^{9}$The College of William \& Mary, Williamsburg, Virginia 23185, USA}
\affiliation{$^{10}$University of Tennessee, Knoxville, Tennessee 37996, USA}
\affiliation{$^{11}$Catholic University of America, Washington, DC 20064, USA}
\affiliation{$^{12}$Hampton University, Hampton, Virginia 23669, USA}
\affiliation{$^{13}$Christopher Newport University, Newport News, Virginia 23606, USA}
\affiliation{$^{14}$Jazan University, Jazan 45142, Saudi Arabia}
\affiliation{$^{15}$Thomas Jefferson National Accelerator Facility, Newport News, Virginia 23606, USA}
\affiliation{$^{16}$Temple University, Philadelphia, Pennsylvania 19122, USA}
\affiliation{$^{17}$Ohio University, Athens, Ohio 45701, USA}
\affiliation{$^{18}$University of Connecticut, Storrs, Connecticut 06269, USA}
\affiliation{$^{19}$Stony Brook University, Stony Brook, New York 11794, USA}
\affiliation{$^{20}$Old Dominion University, Norfolk, Virginia 23529, USA}
\affiliation{$^{21}$University of Colorado Boulder, Boulder, Colorado 80309, USA}
\affiliation{$^{22}$University of New Hampshire, Durham, New Hampshire 03824, USA}
\affiliation{$^{23}$James Madison University, Harrisonburg, Virginia 22807, USA}
\affiliation{$^{24}$Rutgers University, New Brunswick, New Jersey 08854, USA}
\affiliation{$^{25}$Carnegie Mellon University, Pittsburgh, Pennsylvania 15213, USA}