### 2020-09-17 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response/response_email.txt
added PRL response

Sep 16, 2020   8:16 pm

Re: LV16805
    Probing the deuteron at very large internal momenta
    by C. Yero, D. Abrams, Z. Ahmed, et al.

Dear Dr. Yero,

Your manuscript has been reviewed by our referees. We ask you to
consider the enclosed comments.

While we cannot make a definite commitment, we will probably return
your revised manuscript to the previous referees.

Please accompany your resubmittal with a summary of the changes made,
and a brief response to all recommendations and criticisms.

Yours sincerely,

Kevin Dusling
Associate Editor
Physical Review Letters
Email: prl@aps.org
https://journals.aps.org/prl/

NEWS FROM THE PHYSICAL REVIEW JOURNALS

Announcing a new Physical Review journal: PRX Quantum
https://go.aps.org/2Q0uVwP

Celebrating 50 years of Physical Review A, B, C, and D
https://go.aps.org/2IDjVBM

APS is offering promotional open access fee discounts on most
journals in 2020
https://go.aps.org/3388lYG

P.S. We estimate your paper is too long; see appended form.

WORD COUNT

   The guidelines used to determine the word count can be found at
   http://journals.aps.org/authors/length-guide

  Figure   Aspect Ratio   Wide?   Word Equivalent
     1         2.25        Yes        306
     2         0.66         No        247

WORD COUNT SUMMARY
    Note: Text word count excludes title, abstract, byline, PACS,
          receipt date, acknowledgments, and references.

              Text word count   3428 
        Equations word equiv.     16 
          Figures word equiv.    553
           Tables word equiv.      0
                              ------
                       TOTAL    3997 (Maximum length is 3750)

----------------------------------------------------------------------
Report of Referee A -- LV16805/Yero
----------------------------------------------------------------------

See Attachment: lv16805_report_1_a.pdf

----------------------------------------------------------------------
Report of Referee B -- LV16805/Yero
----------------------------------------------------------------------

This letter gives results on the Deuterium dissociation cross section,
and suggests that there may be a transition to a new regime at
transverse momenta of the order of 1 GeV. This is a very interesting
puzzle lying at the border of nuclear physics and QCD. I did not find
any flaw in the analysis presented, and recommend that this letter be
published in PRL.