### 2020-11-17 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response_2/referee_reply_templates.txt
updated prl response 2

===========================
Hi Carlos,

attached are some replies for the referee comments and suggestions for additions/replacements for the introduction.

kind regards

Werner
===========================

1) Understanding the high momentum structure of pn system is highly important for nuclear physics due to the observed 
dominance of short range correlations in nuclei at nucleon momenta above the fermi momentum. This dominance has
been well established by a series of recent experiments carried out at Jefferson Lab and BNL. In these experiments nucleon
momenta up to 1 GeV/c have been probed and recoil momentum distributions have been compared to 
the high momentum part of theoretical deuteron momentum distributions. (cite Piasetky, Korover, Duer and Cruz-Torres)

Recoil momenta up to about 1 GeV/c have also been probed
in a 3He(e,e'p) experiment however at a relatively low momentum transfer of 1.5 (GeV/c)^2 and at a kinematic region (x_b ~1)
where the cross section is dominated by final state interaction. (cite Benmokhtar and Rvachev)

Citations:

E. Piasetzky et al., PRL 97, 162504 (2006)
I. Korover et al, PRL 113, 022501 (2014)
M. Duer et al., PRL 122, 172502 (2019)
R. Cruz-Torre, PRL 124, 212501 (2020)

3He experiments:

F. Benmokhtar, PRL 94, 082305 (2005)
M. M. Rvachev, PRL 94, 192302 (2005)

---------REPLY TO REFEREE COMMENTS---------

2) Our analysis demonstrated that there is no simplistic way of 
estimating relativistic effects since even in the kinematics in which FSI are small 
they are still comparable with the expected relativistic effects. The current calculations
are based on the virtual nucleon approximation in which  relativistic effects are accounted for 
in the propagator of the bound nucleon, for the vertex of electron-bound nucleon interaction as well as
for FSI.  To switch off relativistic effects means to remove all these effects which will return us to the 
standard Glauber theory.  But in our previous paper\cite{2011 PRL paper}  we already demonstrated that Glauber theory 
will not describe the main feature of the reaction. For example at 400 MeV/c it will predict maximum FSI to occur at 90 degrees of 
the spectator angle, while the experiment clearly shows that maximum FSI is at 70 degrees.
The current limitation of the virtual nucleon approximation is that it still uses a non-relativistic deuteron wave 
function(cite Sargsian “private communication”) and assumes that vacuum fluctuation effects are not significant.  It is expected that the latter will grow with the increase 
of the recoil momentum of the spectator nucleon. 
One approach that addresses this issue is the Light-Front approximation, however the implementation of this approximation 
requires a consistent calculation of all the above discussed factors (i.e. propagator, vertex, FSI and wave function).
Hence relativistic effects cannot simply be 'turned off, since they contribute in different components of 
the scattering amplitude.

Light-front calculations are currently underway (M.Sargsian “Private Communication”) which we hope  will be used 
to compare with our data for the large paper.

3) Similar to the above comment regarding relativistic calculations to account for the
effects of different  wave functions, it requires a complete 
calculation including the FSI effects which is out of scope of the current paper. 
We hope that the theory groups who are collaborating with us will be able to provide such 
calculations which will be presented in a larger paper. These studies are currently in progress.

Comment for page 5-6:

The comparison between the measured (reduced) cross sections and the currently available
complete and state of the art calculations illustrates the limit to which a non-relativistic
wave function that is the solution of the Schroedinger equation is valid and able to describe
experimental data that probe the high momentum region of the np system in the most direct way
possible.