### 2021-05-15 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response_2/reply_2/prl_submission_v3/reply_to_RefereeA.txt

--------------------------------
Reply to: Report of Referee A
--------------------------------

A concise, point-by-point response for each of the recommendations and criticisms made by referee A is shown below:

-------------
Referee A:
-------------
1) The connection to SRC is now mentioned. However the references chosen by the authors do not make sense to me.
The standard (general) reference to SRC experimental study is: Subedi et al., Science 320, 1476 (2008). The two
SRC studies highly relevant to this work (emphasizing high momentum) are Korover et al. and A. Schmidt et al.,
Nature volume 578, 540 (2020). Korover et al. is Ref 1 in the new manuscript but the other two are not mentioned.
Moreover, Ref 3 is an odd choice, since it is not directly related to the subject.
----------
Response:
----------
The authors have corrected the references to SRC experimental studies accordingly.

-------------
Referee A:
-------------
2) Relativity (no reference to light-cone formalism), Nucleon-Nucleon interaction (no modern EFT calculations).
I accept the authors response that these theoretical calculations are out of the scope of this experimental paper.
However, to my opinion they should at least be mentioned and discussed in some reasonable way.
----------
Response:
----------
A brief explanation of the light-cone / EFT formalism along with the corresponding references have been added to
the concluding paragraph of the PRL manuscript.

-------------
Referee A:
-------------
3) Their description of why MEC are suppressed is incomplete. The authors give the MEC amplitude, but do not show
that it is smaller than the QE amplitude.
----------
Response:
----------
The description of MEC suppression has been extended to address the QE amplitude.
See page 2 (left column), last paragraph of the PRL manuscript.

-------------
Referee A:
-------------
4) “a weighted average of the cross sections were taken in the overlapping regions of pr”. Show consistency.
They should show the cross section calculated for each setup in the overlap region before they were averaged.
The new manuscript shows the raw data and the corrections in the overlap regions (Supp. Figs 9-11) but not the
cross sections which would allow the reader to verify their consistency and whether the systematic uncertainties
are reasonable.
----------
Response:
----------
The authors have added Fig. 10 in Supplemental Materials to show the reduced cross sections in the overlap
region before being averaged.

-------------
Referee A:
-------------
5) I did not see the requested tables of kinematics and cross sections in the supplemental material.
----------
Response:
----------
The authors did previously include the raw experimental cross-section data (with kinematics)
as well as raw data files with the theoretical cross sections in the form of .txt files.
In addition to the raw data previously submitted, a .pdf version of the data files will also
be submitted in this instance.